### 2025-08-07 paper/paper.tex
add latex stuff

\documentclass[12pt]{amsbook}
\usepackage[utf8]{inputenc}
\usepackage[
backend=biber,
style=numeric,
maxbibnames=99,
sorting=ynt
]{biblatex}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{tikz}
\usepackage{standalone}
\usepackage{svg}
\usepackage{gensymb}
\usepackage{float}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{graphicx, animate}
\usepackage{caption, subcaption}
\usepackage{pgffor}
\usepackage{mathtools}
\usepackage{array}
\usepackage{cleveref}
\addbibresource{bibliography.bib}
\usepackage{hyperref,thmtools}
\linespread{1.5}
\let\cleardoublepage\clearpage
% Override ugly default link
\hypersetup{
  colorlinks   = true, %Colours links instead of ugly boxes
  urlcolor     = blue, %Colour for external hyperlinks
  linkcolor    = blue, %Colour of internal links
  citecolor   = blue    %Colour of citations
}

\begin{document}

\begin{chapter}{Introduction}
    Temporal networks are networks whose nodes and edges change in time. In this thesis we look to model temporal networks. That is, describe the changing temporal network using mathematical language. 
    % Why temp nets are important
    The modelling of temporal networks is an important task in many real world applications including symptom interactions for mental health \cite{jordan2020current,contreras2020temporal}, epidemiology \cite{masuda2013predicting}, and protein interactions \cite{lucas2021inferring,jin2009identifying}.

    % Temp nets are dynam systs
    Temporal networks can be seen as a system in which points, in our case nodes in a network, whose states, the edges connecting them, vary dependent in time, a dynamical system.

    % Hard to find dynam systs because of discrete events 
    Discovering the underlying equations governing these dynamical systems proves challenging. That is because changes in network structure are typically observed in the form of discrete jumps from one state to another; for example an edge between two nodes not being observed at the first time step then being observed at the next.

    % We are proposing a solution
    In this thesis, we propose a hybrid statistical and deep learning framework. That is a framework in which a neural network is used to model probabilistic values of the temporal network. This framework allows us to model temporal networks as continuous-time dynamical systems, discover a fitting set of differential equations describing it, and, exploiting that discovery, predict the time evolution of a temporal network.

    % DEs are useful for modelling interacting states
    Differential equations are useful for modelling systems where the state of one variable can effect the trajectories of other variables. We observe this behavior in temporal networks; nodes' connections within the network can influence the observation of edges between other nodes, for example the phenomenon observed in \cite{newman2001clustering,capocci2006preferential}, where a node is more likely to gain the connections the more connections it has (preferential attachment). With this in mind we might wish to draw on the rich mathematical literature of differential equation modelling.

    % Temp net events are discrete
    In the common representation of networks as binary-valued adjacency matrices, matrices where the observation or lack thereof, of an edge is represented as a 1 or 0 respectively, the events recorded in a temporal sequence of networks correspond to the appearance or the disappearance of link.

    % Difficult to model discrete events
    Because of the discrete nature of events, directly modelling the temporal networks as dynamical systems would require us to handle discrete jumps. The discontinuous character of their temporal evolution, make it challenging to use differential equations techniques.

    % We propose a solution
    Here, we overcome the discreteness problem by interpreting temporal networks as a well established statistical model for complex networks, which embeds nodes in a continuous, low-dimensional metric space by using a truncated singular value decomposition (Random Dot Product Graphs\cite{athreya2017statistical}). In this way we translate the hard problem of modelling discrete events in the space of networks to the easier problem of modelling continuous change in the embedding space.

    % Our solution is interpretable because we use symb reg
    We then define and use Neural Network Differential Equations (NNDE)\cite{SciML_C_Rak} to approximate the time evolution of the embedding space, and symbolic regression techniques to discover the functional form of the fitted NNDEs. These functional forms are interpretable (as they read as classic differential equations) and allow us to predict forward in time the evolution of the temporal networks.

    % We propose a solution
    In this manuscript, we show that the temporal network prediction problem can be successfully re-interpreted as a dynamical system modelling problem by taking the singular value decomposition of a sequence of adjacency matrices and training a NNDE to model an approximation to the underlying differential equation. We then go on to create a functional equation of this approximation.

    % We will do some small tests on the proposed framework
    We apply our proposed framework to three small example temporal networks with the hope of exploring the limitations and strengths of the proposed framework.

    % The framework can and should be tweaked to make it optimised
    The framework we are introducing is extremely flexible, and our research regarding the optimal structure of the Neural Networks used for the NNDEs is just started.
    We are confident that future research can identify more fitting Neural Network structures than the simple one adopted here.
    For this reason, we did not yet attempt to benchmark our model against other classic temporal network prediction methods.

    % The framework can be applied to lots of fields
    As it is completely general, we believe that the framework we are introducing can be usefully applied to areas of medicine, especially protein interaction networks; population dynamics for network ecology; and social network modelling. In particular, we discuss how specific domain knowledge relative to the prediction scenario can be taken into account, moving from NNDEs to Universal Differential Equations (UDEs).
\end{chapter}

\begin{chapter}{Literature Review}

    \section{Temporal Networks}
        % Use of temp nets
        \subsection{Definitions}
        Temporal networks are used to describe a series of interactions across time.
        A temporal network $G$ is defined as:  
        \begin{align}
            G=\{ {_t} G=({_t} N, {_t} E)\} _{t \in 1\ldots T } 
        \end{align}
        $_t x \in {_t} N, {_t} e\in {_t} E$, where ${_t} x, {_t} e$ are nodes and edges at time $t$ respectively, and ${_t} N, {_t} E$ are the sets of nodes and edges at time $t$ respectively.
        An example of this can be found in \cref{method nets}.
        \begin{figure}[H]
            \begin{tabular}{llll}
            \begin{subfigure}[c]{0.3\textwidth}
                \centering
                \resizebox{.6\width}{!}{\input{SynthDataDiagrams/methodology1.tex}}
                \label{method net, a}
            \end{subfigure}
            &
            \centering
            \begin{subfigure}[c]{0.3\textwidth}
                \centering
                \resizebox{.6\width}{!}{\input{SynthDataDiagrams/methodology2.tex}}
                \label{method net, b}
            \end{subfigure}
            &
            $\cdots$
            &
            \centering
            \begin{subfigure}[c]{0.3\textwidth}
                \centering
                \resizebox{.6\width}{!}{\input{SynthDataDiagrams/methodology3.tex}}
                \label{method net, c}
            \end{subfigure}

            \end{tabular}
            \caption{Example sequence of networks.}
            \label{method nets}

        \end{figure}
        %representations of temp nets as adj lists
        This can be represented in many ways, for example as a series of interactions. The series of interactions is generally represented as a list of tuples containing the nodes which interacted and the time that the interaction took place. 
        \begin{align}
            G=\{(i, j, t)|\space i,j\in {_t} N, (i\leftrightarrow j)\in {_t}E \}
        \end{align}

        % representation of temp nets as adj mats
        A temporal network can also be represented as a sequence of adjacency matrices. In the representation of a series of adjacency matrices, the columns and rows of the matrix each refer to each of the nodes. In this way, if an edge exists between nodes $i$ and $j$, the entry $(i,j)$ of the adjacency matrix will be 1; if an edge does not exist the entry will be 0.
        \begin{align}
            _t G &= {_t} A\\
            _t A_{ij} &=  
            \left\{
                \begin{array}{ll} 
                1 & (i\leftrightarrow j)\in {_t} E\\
                0 & otherwise
                \end{array}
            \right.
        \end{align}
        The row $_t A_{i\cdot}$ describes the outgoing edges of the node $i$, while the column $_t A_{\cdot j}$ describes the incoming edges of the node $j$.

        \subsection{Applications}
        Temporal networks are present in many areas of research interest, ranging from symptom interactions in mental health\cite{jordan2020current,contreras2020temporal}, to epidemiology\cite{masuda2013predicting}, protein interactions\cite{lucas2021inferring,jin2009identifying}, and social networks\cite{moinet2015burstiness,hanneke2010discrete}.

        % Overview of modelling attempts
        Generally attempts to model temporal networks either model the change in node state or the change in network structure.

        The node state describes some information about that node:
        \begin{align}
            _t x &= (m_1,\ldots, m_p)
        \end{align}
        Where $m_i$ is some metadata about the node $_t x$.

        An example of modelling the change in a node state could be the severity of a symptom at a given time, as in \Citeauthor{contreras2020temporal}\cite{contreras2020temporal}. Where the node $_t x$ represents a given symptom and $m_1$ would represent the severity of that symptom.

        % Intro contreras as attempt to model changes in nodes
        \Citeauthor{contreras2020temporal}\cite{contreras2020temporal} model the change in node state over time while the structure of the network remains constant. The authors look at the severity of symptoms over time, i.e. the state of the nodes, where the nodes exist in a statically structured temporal network.

        % Contreras methods
        \Citeauthor{contreras2020temporal} use multilevel vector autoregression, a method that finds correlations between the current state of a variable and the states of variables at previous time steps \cite{singer2003applied}, to model how the current symptoms of people with paranoia might predict their future symptoms. These models were then used to create three networks that linked symptoms.

        % Focal point of the paper is their temporal network
        Of particular interest to this thesis is the temporal network created by having a fully connected network of all symptoms including self-loops, where a symptom connects to itself; this self loop comes about from a symptom being correlated with itself from one time step to the next. The edge weights represent the extent to which the severity of each symptom at time $(t)$ predicts the severity of itself and other symptoms at time $(t+1)$.

        % Why this is the focal point
        This framework keeps the overall structure of the network fixed throughout time and so is not particularly flexible, but is extremely human interpretable and can be used to extrapolate and predict into the future.

        % intro to paper as attempt to model node state
        In \Citeauthor{KARIMI20133476}\cite{KARIMI20133476}, \Citeauthor{KARIMI20133476} look to model the change in node state while the network structure changes.

        % How network is represented
        \Citeauthor{KARIMI20133476} represent their temporal network as a series of interactions (a list of tuples in the form $(i,j,t)$, where $i,j$ are the nodes interacting and $t$ is the time the at which the interaction occurred).

        % what state the paper attempts to model
        The authors use a model to predict when a node will change from state 0 to state 1, in practice this could represent whether a user of a social network believes a rumour, or whether an ecological patch has been colonised.

        % Model limitation
        Notably, in this model, the node never changes back; this is done to keep the model analytically tractable, but does limit its usefulness in some contexts. In this model an organism could not become locally extinct in an ecological patch for example.

        % Methods to model node state
        To model the change in a nodes' state, the model uses a sliding time window and if the fraction of interactions with nodes of state 1 within that time window exceeds a given threshold, the node switches to state 1 as well.

        % Limitation of model
        In this work we aim to present a model for predicting the state of a node in the future. Instead, \Citeauthor{KARIMI20133476} aim to present a tool for understanding the spread of ideas, viruses, etc. within a network.

        Other efforts focus on modelling the changing structure of the network as in \Citeauthor{sanna2021link}\cite{sanna2021link}, which looks to model $_t G$ as a whole.

        % temp net representation
        In this case \Citeauthor{sanna2021link} represent the temporal network as a series of adjacency matrices, each matrix representing an observation of the changing network.

        % Methods for embedding
        The authors look to model the structure of edges by using a spectral embedding (SVD). A spectral embedding uses the eigenvectors of a network's adjacency matrix to find vectors that represent the nodes of the network. The spectral embedding transforms their temporal sequence into a sequence of observations of a latent space.

        % What is a latent space
        A latent space is a continuous space in which similar nodes (nodes with many of the same neighbours) are positioned near each other. In this space nodes are represented as points in a continuous, low dimensional space. Another feature of this latent space, is that the dot product of any two points in the space gives us the probability that the nodes related to those points are connected, a Random Dot Product Graph (RDPG)\cite{athreya2017statistical}.

        % methods for time series
        With this series of embeddings, the authors employ a variety of time series techniques to predict the network structure at future time steps.

    \section{UDEs and NNDEs}
        % Justification for DEs being useful
        In some temporal networks the state of a node might influence the evolution of its neighbours. For example, in social networks we often observe that the more neighbours a node has, the more likely they are to gain more neighbours, a power law distribution of edges\cite{zhao2012multi,garg2009evolution}. We also see that the state of symptoms (occurrence, severity, and distress), can influence the states of other symptoms in patients undergoing chemotherapy \cite{papachristou2019network,kalantari2022network}, and for mental health disorders \cite{contreras2020temporal}. As well as in ecological networks where the interactions of species are often represented as a network, and the populations of one species may influence the population growth of another \cite{elton2001animal,volterra1927variazioni}; that is, the state (in this case the population) of a species (node), may influence the movement of another species.

        % DEs have lots of research
        With the observation that nodes in real world temporal networks may influence others, differential equations or difference equations seem appropriate for modelling the evolution of nodes; there is a rich body of literature exploring these methods use in modelling interacting variables.

        % The data we have is discrete, which is a problem for DEs
        One of the issues that the differential equation method encounters is the prevalence of discrete jumps of edges being observed and then not being observed. Because of this, we may consider exploring the possibility of using difference equations.

        % time is continuous so good to have our model be cont too
        We note that the progression of the network is generally continuous even if the events happen as discrete jumps. We may wish to preserve this continuous temporal progression in our model.

        %This approach may overcome the problem of discrete observations, by defining a progression function that will only predict discrete formation or decay of edges\cite{hanneke2010discrete}. However, this imposes a significant limitation on the temporal granularity with which we can predict the state of the network. We would be limited to the granularity our data is collected at. This may be a problem if for example we had a series of weekly or monthly observations of networks we would not be able to predict interactions at each day. We can overcome both this limitation and the discrete jump limitation by using random dot product graphs. Interpolation may be possible in this case, but we would have no way of knowing if the interpolation had any basis in the real world.

        % passino has discrete data but uses a method to view it as cont
        Looking at the work of \Citeauthor{sanna2021link}\cite{sanna2021link}, the authors use RDPGs to approximate a temporal network as a matrix of probabilities of an edge existing between two nodes. In contrast to edges, these probabilities can evolve continuously, and so we can use differential equations to generate probabilistic networks at any temporal resolution we require. That is, we will be able to model the changes in the probabilities of edges continuously and from this, generate a network at any time step.

        %  we can use the same method we can view data as dynam sys and use DEs
        Using the method in \Citeauthor{sanna2021link}\cite{sanna2021link}, we can treat the sequence of embeddings as a dynamical system and then use differential equation modelling in a continuous space. Once we have found a latent space for each network, we know that each point in this space refers to a specific node in the network, and so, as the network changes, so too does the position of each of the points in the latent space. We then look to model this movement using differential equations.

        % Can use a new NN structure to make this framework very general
        Our aim for this framework is to be as general as possible and note that there may not yet be enough domain knowledge to perform traditional differential equation modelling to the desired accuracy. For example in network ecology models can be useful, but not perfect. Hence, we might wish to build on these models rather than trying to create something entirely new. This can be achieved by using this partial domain knowledge using it to create a UDE\cite{SciML_C_Rak}.

        \subsection{Theory}

            % Overview of why UDEs are useful
            Developed by \Citeauthor{SciML_C_Rak}\cite{SciML_C_Rak}, UDEs are a novel neural network architecture that aims to take the best of both traditional neural differential equation modelling and of flexible machine learning approaches to modelling.

            If we consider a comparison to an Ordinary Differential Equation (ODE), the ODE is written as:
            \begin{align}
                y(0) &= y_0 \\
                \frac{dy}{dt}(t)&=f(t,y(t))
            \end{align}

            The UDE will incorporate some component of $f(t,y(t))\approx \hat f(t,y(t))$, and will learn the difference with a Neural Network (NN) such that:
            \begin{align}
                \label{UDE}
                y(0) &= y_0 \\
                \frac{dy}{dt}(t)&=\hat f(t,y(t))+NN(t,y(t), \theta)
            \end{align}
            Where $\theta$ is a vector of the parameters of the NN.
            % Structure of UDE
            UDEs combine domain knowledge in the form of a differential equation, with a NN. This hybrid machine learning model is then trained on the observed data.

            % Overview of why UDEs are useful
            Combining the two allows for most of the movement in the data to be captured by the differential equation, which contains the domain knowledge. In this way, the NN will only need to learn a theoretically simpler equation, given that a part of the system has already been captured.

            % Why this structure is useful
            This then allows the NN to be smaller, require fewer data, and be trained much faster than traditional NN models. Given the flexibility of NNs, the use of UDEs is natural when modelling a process that is not entirely understood\cite{kidger2022neural}.

            Note that we can construct a special case of a UDE, an NNDE when $\hat f(t,y(t))=0$. This structure can be used when do not have any domain knowledge we wish to give the model.

            % Software used for coding
            There is an ecosystem of packages specifically designed for performing machine learning with UDEs. The ecosystem is called Scientific Machine Learning (SciML)\cite{SciML_C_Rak}. Largely, the SciML ecosystem has been optimised for flexibility and efficiency with respect to models available and training performance. Given there does not seem to be any other ecosystem with this feature, it seemed like 
            an obvious choice.

        \subsection{Applications}
            Whilst there has been much interest in implementing UDEs and NNDEs, much of this work has been focussed on physics informed neural network and physics informed neural ordinary differential equations (PINNs and PINODEs) \cite{karniadakis2021physics,GAO2021110079,krishnapriyan2021characterizing,roehrl2020modeling}, and on improving modelling of fluids \cite{mahmoudabadbozchelou2021data,nguyen2022physics}. Alongside this, UDEs have gained popularity in the last few years, and there have been a number of studies exploring their usefulness in modeling the effect of restrictions due to COVID-19 on the virus' spread \cite{Dandekar2020.04.03.20052084}. Although there has been a large amount of research into the usefulness of these types of models, to the best of this author's knowledge they have never been applied to the problem of predicting temporal networks. This seems to be a natural fit for UDEs and NNDEs, given we may know relatively little about the processes that govern the evolution of temporal networks, and any information we do have can be incorporated into the model.

    \section{Symbolic Regression}
        % What is symbolic regression
        Symbolic regression is a type of analysis that looks to take input variables and find a relationship between them that describes the output variables. For symbolic regression, we construct a set of mathematical operations that may be used to find this relationship. The algorithm then generates combinations of these equations to find an expression that matches the output variables. Because we construct the set of potential operations, the resulting expression is interpretable by humans (a functional equation), as opposed to a NN for example (a numeric equation).

        % Why is symbolic regression useful
        Having interpretable models that capture underlying relations is relevant to all fields of study, and so the applications of symbolic regression are incredibly varied.

        % Where has symbolic regression been used 
        Various methods of symbolic regression have been use in manufacturing systems, chemical systems, and tumor research\cite{can2011comparison,keith2021combining,yoshihara2013inferring}. As well research into improvements is still on going.

        % How is this done
        Finding a suitable functional expression can be achieved in many ways and remains active area of research. One of the most popular algorithms is genetic programming\cite{schmidt2009distilling}.

        % How genetic programming works
        Genetic programming involves constructing a search tree using an evolutionary algorithm to yield candidate equations. An evolutionary algorithm is an iterative process. It begins by randomly generating a set of candidate equations with operations from the given set of operations:
        \begin{align}
            Q_1=\{f_1,\dots,f_n\}
        \end{align}
        Each candidate function $f_i$ is tested and given a fitness score based on how well it performed. The higher a candidates' fitness, the more likely it is to be selected for the next iteration.

        Suppose $f_1,f_2$ were selected for the next iteration. The evolutionary algorithm would then randomly generate new functions based on $f_1,f_2$ again with the given set of operations.
        \begin{align}
            Q_2=\{f_{1,1},\dots,f_{1,n}, f_{2,1},\dots,f_{2,n}\}
        \end{align}
        This process is repeated until a suitable functional equation is found.

        For the genetic programming algorithm, the fitness of a candidate function is determined by how closely it matches the data as well as numerically calculated partial derivatives of each of the input variables.

        % Why are the derivatives important
        Comparing with these derivatives is useful as it helps to ensure that the produced function equations have some grounding in the real world process that generated the data. The search tree is then pruned by comparing the partial derivatives of the candidate equation, with partial derivatives calculated from data. This process is then iterated many times to generate a suitable equation.

        % What symbolic regression aims to solve
        The problem that symbolic regression aims to solve is to generate interpretable and ``meaningful'' equations from observed data.

        % Examples of use of genetic programming
        For example, both \Citeauthor{schmidt2009distilling}\cite{schmidt2009distilling}, and \Citeauthor{bongard2007automated}\cite{bongard2007automated} test their methods' capacity for obtaining a functional equation with physical simulation data, various pendula specifically. A ``meaningful'' equation in this case may be an equation that conserves momentum.

        % Expalaination of schmidt
        Examples of genetic programming include discovering equations governing the movement of simple harmonic and chaotic double pendula from data. In \Citeauthor{schmidt2009distilling}\cite{schmidt2009distilling}, the authors present an evolutionary algorithm to generate multiple candidate functions that approximate the partial derivatives of each variable that are calculated directly from the data.

        % Intro to bongard
        Another approach to symbolic regression can be found in \Citeauthor{bongard2007automated}\cite{bongard2007automated}. The authors propose a method to find a functional equation that approximates a learned numerical equation rather than directly from data.

        % How does this process work
        The algorithm generates a set of candidate equations. The candidate equations are then tested against the numerical equation in such a way that the difference between the candidate and numerical predictions are maximised. The algorithm will generate new initial conditions to find behaviour exhibited by the numerical equation, for example a fixed point, that is not exhibited by the functional equation. When a difference is found, the functional equations are adjusted to also exhibit the behaviour of the numerical equation.

        % Limitations of this process
        We will be using a UDE with relatively small data sets. The models will be trained on a small range of input values, and so we would not have confidence that any behaviour exhibited outside the range of these input values would reflect the underlying system. As such, this method may not yield us useful results.

        % Benefits of combining UDEs and symbolic regression
        In very recent work, Kidger\cite{kidger2022neural} demonstrates that, with UDEs, two of the assumptions necessary for symbolic regression can be overcome. Those are the requirement for paired observations and derivatives, as well as the assumption that the function can be expressed as a shallow tree of symbolic operations. We are modelling a differential equation, and so will have access to the derivatives of our numerical model at any point, as neural networks are easily differentiable.

        % Usefulness of symbolic regression for nnde
        In \Citeauthor{SciML_C_Rak}\cite{SciML_C_Rak}, the authors demonstrate that by using symbolic regression to find a functional equation of our model, as opposed to just a UDE, we can improve the accuracy of extrapolating.
\end{chapter}

\begin{chapter}{Methodology}

        % Goal is to predict next time steps
        We are given a sequence of graphs, and our goal is to predict the next graph in the sequence.

        % How we achieve this goal
        To achieve this, we reinterpret our temporal network as a sequence of points in a latent space; that is, a space in which similar vertices (ones with many of the same neighbours) are mapped to a similar point in space.
        We then train a Neural Network Differential Equation (NNDE) to approximate the rate of change of the points in this latent space. With this numerical equation we can then use symbolic regression techniques to find a functional equation that matches the numerical equation. Using this process we discover an interpretable function that governs the evolution of a temporal network.

        \begin{figure}{H}
            \centering

            \begin{subfigure}[c]{1\textwidth}
                \begin{tabular}{llll}
                    $\begin{bmatrix}
                        \cdot & 1 & 1 & 1 & \cdot & \cdot & \cdot & \cdot\\
                        1 & \cdot & 1 & \cdot & \cdot & \cdot & \cdot & \cdot\\
                        1 & 1 & \cdot & 1 & 1 & \cdot & \cdot & \cdot\\
                        1 & \cdot & 1 & \cdot & \cdot & \cdot & \cdot & \cdot\\
                        \cdot & \cdot & 1 & \cdot & \cdot & 1 & 1 & \cdot\\
                        \cdot & \cdot & \cdot & \cdot & 1 & \cdot & \cdot & \cdot\\
                        \cdot & \cdot & \cdot & \cdot & 1 & \cdot & \cdot & 1\\
                        \cdot & \cdot & \cdot & \cdot & \cdot & \cdot & 1 & \cdot
                    \end{bmatrix}$
                    &
                    $\begin{bmatrix}
                        \cdot & 1 & 1 & 1 & \cdot & \cdot & \cdot & \cdot\\
                        1 & \cdot & 1 & 1 & \cdot & \cdot & \cdot & \cdot\\
                        1 & 1 & \cdot & 1 & 1 & \cdot & \cdot & \cdot\\
                        1 & 1 & 1 & \cdot & \cdot & \cdot & \cdot & \cdot\\
                        \cdot & \cdot & 1 & \cdot & \cdot & 1 & 1 & \cdot\\
                        \cdot & \cdot & \cdot & \cdot & 1 & \cdot & \cdot & 1\\
                        \cdot & \cdot & \cdot & \cdot & 1 & \cdot & \cdot & 1\\
                        \cdot & \cdot & \cdot & \cdot & \cdot & 1 & 1 & \cdot
                    \end{bmatrix}$
                    &
                    $\cdots$
                    &
                    $\begin{bmatrix}
                        \cdot & 1 & 1 & 1 & \cdot & \cdot & \cdot & \cdot\\
                        1 & \cdot & 1 & 1 & \cdot & \cdot & \cdot & \cdot\\
                        1 & 1 & \cdot & 1 & \cdot & \cdot & \cdot & \cdot\\
                        1 & 1 & 1 & \cdot & \cdot & \cdot & \cdot & \cdot\\
                        \cdot & \cdot & \cdot & \cdot & \cdot & 1 & 1 & \cdot\\
                        \cdot & \cdot & \cdot & \cdot & 1 & \cdot & 1 & 1\\
                        \cdot & \cdot & \cdot & \cdot & 1 & 1 & \cdot & 1\\
                        \cdot & \cdot & \cdot & \cdot & \cdot & 1 & 1 & \cdot
                    \end{bmatrix}$

                \end{tabular}
                \caption{Sequence of sparse adjacency matrices associated with the networks in \cref{method nets}.}
                \label{method adjacency}
            \end{subfigure}

            \begin{subfigure}[c]{1\textwidth}
                \begin{tabular}{llll}
                    $\begin{bmatrix}
                        -0.82  & 0.01\\
                        -0.64  & 0.36\\
                        -0.95 & -0.76\\
                        -0.64 &  0.36\\
                        -0.48 &  0.82\\
                        -0.17 & -0.40\\
                        -0.20 & -0.52\\
                        -0.07 &  0.25\\
                    \end{bmatrix}$

                &

                    $\begin{bmatrix}
                        -0.82  &  0.10\\
                        -0.82  &  0.10\\
                        -0.90 & -0.58\\
                        -0.82  &  0.10\\
                        -0.42 &  0.81\\
                        -0.13  & -0.64\\
                        -0.13  & -0.64\\
                        -0.12 &  0.57\\
                    \end{bmatrix}$

                &
                $\cdots$
                &

                    $\begin{bmatrix}
                        -0.85 &  -0.10\\
                        -0.85 &  -0.10\\
                        -0.85 &  -0.10\\
                        -0.85 &  -0.10\\
                         0.08 &  -0.69\\
                         0.11 &  -0.88\\
                         0.11 &  -0.88\\
                         0.08 &  -0.69\\
                    \end{bmatrix}$
                \end{tabular}
                \caption{Sequence of truncated SVDs associated with the networks in \cref{method nets}.}
                \label{method embed}
            \end{subfigure}

            \begin{subfigure}[c]{1\textwidth}
                \begin{tabular}{llll}
                \begin{subfigure}[c]{0.31\textwidth}
                    \centering
                    \begin{subfigure}[p]{1\textwidth}
                        \includegraphics[width=\linewidth]{../Code/Plots/examples/plot 1.png}

                    \end{subfigure}

                \end{subfigure}
                &
                \centering
                \begin{subfigure}[c]{0.31\textwidth}
                    \centering
                    \begin{subfigure}[p]{1\textwidth}
                        \includegraphics[width=\linewidth]{../Code/Plots/examples/plot 2.png}

                    \end{subfigure}
                \end{subfigure}
                &
                $\cdots$
                &
                \centering
                \begin{subfigure}[c]{0.31\textwidth}
                    \centering
                    \begin{subfigure}[p]{1\textwidth}
                        \includegraphics[width=\linewidth]{../Code/Plots/examples/plot 3.png}

                    \end{subfigure}
                \end{subfigure}

                \end{tabular}
                \caption{Sequence of embedded points associated with the networks in \cref{method nets}.}
                \label{method plots}

            \end{subfigure}

            \caption{Visual illustration of the proposed framework.}
            \label{framework illustration}
        \end{figure}

    \section{Singular Value Decomposition and Random Dot Product Graphs}
            \label{svd}
            % Over view of section
            In this section we discuss the Singular Value Decomposition (SVD) of a matrix, and its usefulness in creating Random Dot Product Graphs (RDPG).

            % Definition of SVD
            We define the matrix $A$ to be a real $m \times n$ with $n \le m$. $A$ can be expressed as \cite{forsythe1967computer}:
            \begin{align}
                A&=l\sqrt{\Sigma} \sqrt{\Sigma} r' \\
                A&=L R' \quad \text{where} \; L=l\sqrt \Sigma,\: R=\sqrt \Sigma r'
            \end{align}
            Where $l, r$ are real valued, orthonormal matrices. The columns of $l$ consist of the $n$ largest eigenvectors of $AA'$ and the columns $r$ consist of the eigenvectors of $A'A$. $\Sigma$ is a diagonal matrix whose entries are the square root of the positive eigenvalues of $A'A$ in decreasing order.

            % How to reduce problem size
            In order to reduce the problem size, we can truncate the $L,R$ to the first $d$ columns to yield, $\hat L, \hat R$ such that:

            \begin{align}
                A &\approx \hat L \hat R' \\
                \hat A &= \hat L \hat R'
                \label{approx}
            \end{align}

            % Supposed to be how to interpret this as RDPG but is bad
            The approximation of $A$ is defined as $\hat A$ and is given by \cref{approx} can be interpreted as a graph where the entry $(i,j)$ of the matrix is the probability that an edge $(i \leftrightarrow j)$ exists. This approximation $\hat A$ is an RDPG\cite{athreya2017statistical}.

            Using an RDPG lets us model the probabilities of edges existing as continuous variables, as opposed to the discontinuous observations themselves. A benefit of using an SVD to create the RDPG, is that, to model the probabilities of $\hat A$, we can model the change of $\hat L$ and $\hat R$ separately. This is useful because the rows of the matrices $\hat L$ and $\hat R$ can be viewed as vector points in a $d$-dimensional space, where $d$ is the number of singular values we take for our SVD truncation. Viewing $\hat L$ and $\hat R$ as a collection of vector points allows us to change the problem of modelling changing matrices to a problem of modelling a dynamical system. This modelling can be accomplished by finding an appropriate DE.

            % SVD can give a latent space embedding
            The SVD can be considered to give a latent space representation of the matrix\cite{hoff2002latent}. That is, in this space similar nodes in the network are mapped to similar points in space. A visual representation of this process can be seen in \cref{framework illustration}.

            % Introduce problem of alignment
            It is important to note that the SVD is not unique. The distances between points will be the same for every embedding, but the orientation may be different, and so, if the SVD of each time step is not aligned, learning will be effectively impossible as we will essentially be trying to model random noise.

            % How to solve alignment problem
            Here, we can use Procrustes alignment\cite{DotProductGraphs} to align the matrices after they have been generated.

            % Explicit process for how we process our data
            When obtaining the RDPGs for our temporal network, we take each network observation \cref{method nets} and represent them in the form of an adjacency matrix \cref{method adjacency}. We then use a truncated singular value decomposition to reduce the dimensionality of our adjacency matrices\cite{golub1971singular}.

            % Introduce problem of loss from SVD
            This truncation introduces some loss, but can dramatically reduce the size of our problem. \Citeauthor{Runghen2021}\cite{Runghen2021} effectively employ this method to reduce a problem involving over 130,000 visitors to a 6 dimensional latent space, whilst preserving 70\% of the information.

            % What code we use
            To execute the SVD we use the ARPACK Julia package \cite{lehoucq1998arpack}. This package uses an iterative approach to approximate the eigenvectors and singular values of each of our adjacency matrices\cite{lehoucq1996deflation}.

    \section{Neural Network Differential Equation}
        Once we have the SVD embedding of points, we look to model the trajectories of a target point. That is, with some input data, point location, distance to neighbours, etc, we want to model the differential equation that governs the movement of a point. 
        % Modelling multiple points
        Notably this method can be used to model multiple points by simply changing the node being predicted and trained on.

        From \cref{UDE} our problem can be represented as the UDE:
        \begin{align}
            \frac{d \hat A}{dt} &= \hat f(t,\hat A(t)) + NN(t, \hat A, \theta)
        \end{align}
        The focus of this thesis is to understand the behaviour of the neural network when modelling temporal neural networks. Hence, we set $\hat f(t, \hat A(t))=0 \space \forall t$. With this, the problem becomes:
        \begin{align}
            \frac{d \hat A}{dt} &= NN(t, \hat A, \theta)
        \end{align}
        A special case of a UDE called a Neural Network Differential Equation(NNDE).

        In this thesis, as a proof of concept, we focus on a modelling the change in structure of a single node, referred to as the target node $u$, where $u(t)={_t}\hat L_{1\cdot}$. To model $u$ we use the direction vectors from $u$ to its $k$ (15) nearest neighbours in $\hat L$. We define the function $K(\hat L(t), k)$ to find the direction vectors of the $k$ nearest neighbours to $u(t)$. We use the function $K$ to limit the complexity of our model.

        \begin{align}
            \frac{du}{dt} &= NN(t, K(\hat L, k), \theta)
        \end{align}

>>>>

\documentclass[12pt]{amsbook}
\usepackage[utf8]{inputenc}
\usepackage[
backend=biber,
style=numeric,
maxbibnames=99,
sorting=ynt
]{biblatex}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{tikz}
\usepackage{standalone}
\usepackage{svg}
\usepackage{gensymb}
\usepackage{float}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{graphicx, animate}
\usepackage{caption, subcaption}
\usepackage{pgffor}
\usepackage{mathtools}
\usepackage{array}
\usepackage{cleveref}
\addbibresource{bibliography.bib}
\usepackage{hyperref,thmtools}
%\usepackage{chngcounter}

\usetikzlibrary{decorations.markings}
\usetikzlibrary{arrows,automata}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta,positioning}

\counterwithin{figure}{chapter}
\counterwithin{section}{chapter}
\counterwithin{table}{chapter}
\counterwithin{subsection}{section}

\linespread{1.5}
\let\cleardoublepage\clearpage
% Override ugly default link
\hypersetup{
  colorlinks   = true, %Colours links instead of ugly boxes
  urlcolor     = blue, %Colour for external hyperlinks
  linkcolor    = blue, %Colour of internal links
  citecolor   = blue    %Colour of citations
}

\begin{document}
\begin{chapter}{Methodology}

        % Goal is to predict next time steps
        We are given a sequence of graphs, and our goal is to predict the next graph in the sequence. To achieve this, we reinterpret our temporal network as a sequence of points in a latent space; that is, a space in which similar vertices (ones with many of the same neighbours) are mapped to a similar point in space.
        We then train a Neural Ordinary Differential Equation (NODE) to approximate the rate of change of the points in this latent space. With this numerical equation, we can then use symbolic regression techniques to find a functional equation that matches the numerical equation. Using this process we discover an interpretable function that governs the evolution of a temporal network.

    \begin{        Initial vector is [1 2 3 4 ... mod 7] (Double check)
        Initial vector is used to align the vectors but it kind of doesn't work
section}{Singular Value Decomposition and Random Dot Product Graphs}
        SVD Calculated with XXX Arpack XXX Julia
        Each system has 2 singular values
        L and R are related by [1 0; 0 -1] as the network is not directed (IE the matrix is symetrical) 
        We only model L and transform it for analysis
        Reconstruct the RDPG L*(L*transform)'
    \end{section}

    \begin{section}{Neural Ordinary Differential Equation}
        Used Julia SciML ecosystem
        Basic copy pasted code 
        Modelling the whole state
        Loss function is abs2 sum Data-pred + disabove 1 and below 0 (L*(L*transform)' )
        NN structure 3 layers 256, 128, 128 celu activation function
    \end{section}
